Proving linearizability using forward simulations\thanks{An extended version of this paper including the missing proofs can be found at \cite{extended}.}

### 2017-05-05 draftCAV/queue.tex
first part of H&Q proof

%An enqueue operation in a $\mathit{HWQ}$ state is pending and present if and only if its argument is stored in the array ${\tt items}$ or it has not yet written to the array  ${\tt items}$. Those operations should be $\mathsf{pending}$ and $\mathsf{present}$ in the related $AbsQ$ states. In addition to this, the forward simulation $F_1$ imposes following restrictions:
An $\mathit{HWQ}$ state $s$ is related by $F_1$ to $AbsQ$ states $t$ where the predicate
$\mathsf{present}$ is true for all the enqueues in $s$ whose argument is stored in the array ${\tt items}$, and all the pending enqueues that have not yet written to the array ${\tt items}$ (and only for these enqueues). We refer to such enqueues in $s$ as $\mathsf{present}$ enqueues. Also, $\mathsf{pending}(k)$ is ${\tt true}$ in $t$ whenever $k$ is a pending enqueue in $s$, $\mathsf{arg}(d)=k$ in $t$ whenever the argument of the enqueue $k$ in $s$ is $d$, and for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that ${\tt y}(k)=d$ (recall that ${\tt y}$ is a local variable of the dequeue method in $AbsQ$). The order relation $\mathsf{before}$ in $t$ satisfies the following constraints:
\vspace{-2mm}
\begin{itemize}
	\item[(a)] pending enqueues are maximal, i.e., for every two $\mathsf{present}$ enqueues $k$ and $k'$ such that $k'$ is $\mathsf{pending}$, we have that $\neg \mathsf{before}(k',k)$, % $k'\not< k$,
	\item[(b)] $\mathsf{before}$ is consistent with the order in which positions of ${\tt items}$ have been reserved, i.e., for every two present enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $\neg \mathsf{before}(k',k)$, %$k' \not< k$,
	\item[(c)] if the position $i$ reserved by an enqueue $k$ has been ``observed'' by a non-linearized dequeue that in the current array traversal may ``observe'' a later position $j$ reserved by another enqueue $k'$, then $k$ can't be ordered before $k'$,
%	an enqueue which has reserved a position $i$ %and executed only the first statement
%	can't be ordered before another enqueue that has reserved a position $j \geq i$ ,
	i.e., for every two present enqueues $k$ and $k'$, and a dequeue $k_d$, such that

>>>>

%An enqueue operation in a $\mathit{HWQ}$ state is pending and present if and only if its argument is stored in the array ${\tt items}$ or it has not yet written to the array  ${\tt items}$. Those operations should be $\mathsf{pending}$ and $\mathsf{present}$ in the related $AbsQ$ states. In addition to this, the forward simulation $F_1$ imposes following restrictions:
An $\mathit{HWQ}$ state $s$ is related by $F_1$ to $AbsQ$ states $t$ where the predicate
$\mathsf{present}$ is true for all the enqueues in $s$ whose argument is stored in the array ${\tt items}$, and all the pending enqueues that have not yet written to the array ${\tt items}$ (and only for these enqueues). We refer to such enqueues in $s$ as $\mathsf{present}$ enqueues. Also, $\mathsf{pending}(k)$ is ${\tt true}$ in $t$ whenever $k$ is a pending enqueue in $s$, $\mathsf{arg}(k)=d$ in $t$ whenever the argument of the enqueue $k$ in $s$ is $d$, and for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that ${\tt y}(k)=d$ (recall that ${\tt y}$ is a local variable of the dequeue method in $AbsQ$). The order relation $\mathsf{before}$ in $t$ satisfies the following constraints:
\vspace{-2mm}
\begin{itemize}
	\item[(a)] pending enqueues are maximal, i.e., for every two $\mathsf{present}$ enqueues $k$ and $k'$ such that $k'$ is $\mathsf{pending}$, we have that $\neg \mathsf{before}(k',k)$, % $k'\not< k$,
	\item[(b)] $\mathsf{before}$ is consistent with the order in which positions of ${\tt items}$ have been reserved, i.e., for every two present enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $\neg \mathsf{before}(k',k)$, %$k' \not< k$,
	\item[(c)] if the position $i$ reserved by an enqueue $k$ has been ``observed'' by a non-linearized dequeue that in the current array traversal may ``observe'' a later position $j$ reserved by another enqueue $k'$, then $k$ can't be ordered before $k'$,
%	an enqueue which has reserved a position $i$ %and executed only the first statement
%	can't be ordered before another enqueue that has reserved a position $j \geq i$ ,
	i.e., for every two present enqueues $k$ and $k'$, and a dequeue $k_d$, such that
